---
title: "Credit Risk Assessment"
author: "Edinor Junior"
date: "Feb 08, 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project - Credit Risk Assessment

For this analysis, we will use a set of German Credit Data, already clean and organized for the creation of the predictive model.

The entire project will be described according to its stages.

This project is a initial study about credit risk assessment, build with the intent to create a base to more elaborate works. 

Please, feel free to contribute and share your knowledge

## Step 1 - Collecting the Data

Here is the data collection, in this case a csv file.

```{r collect}
# Data Collecting
credit.df <- read.csv("credit_dataset.csv", header = TRUE, sep = ",")
```

## Step 2 - Normalizing the Data

```{r normalizing}
## Converting variables to type factor (categorical)
to.factors <- function(df, variables){
  for (variable in variables){
    df[[variable]] <- as.factor(df[[variable]])
  }
  return(df)
}

## Normalizing
scale.features <- function(df, variables){
  for (variable in variables){
    df[[variable]] <- scale(df[[variable]], center=T, scale=T)
  }
  return(df)
}

# Normalizing variables
numeric.vars <- c("credit.duration.months", "age", "credit.amount")
credit.df <- scale.features(credit.df, numeric.vars)

# Variables of type factor
categorical.vars <- c('credit.rating', 'account.balance', 'previous.credit.payment.status',
                      'credit.purpose', 'savings', 'employment.duration', 'installment.rate',
                      'marital.status', 'guarantor', 'residence.duration', 'current.assets',
                      'other.credits', 'apartment.type', 'bank.credits', 'occupation', 
                      'dependents', 'telephone', 'foreign.worker')

credit.df <- to.factors(df = credit.df, variables = categorical.vars)

```

## Step 3 - Splitting the data into training and test data

```{r training}
# Dividing the data into training and testing - 60:40 ratio
indexes <- sample(1:nrow(credit.df), size = 0.6 * nrow(credit.df))
train.data <- credit.df[indexes,]
test.data <- credit.df[-indexes,]

```

## Step 4 - Feature Selection

```{r performance}
library(caret) 
library(randomForest) 

# Function to select variables
run.feature.selection <- function(num.iters=20, feature.vars, class.var){
  set.seed(10)
  variable.sizes <- 1:10
  control <- rfeControl(functions = rfFuncs, method = "cv", 
                        verbose = FALSE, returnResamp = "all", 
                        number = num.iters)
  results.rfe <- rfe(x = feature.vars, y = class.var, 
                     sizes = variable.sizes, 
                     rfeControl = control)
  return(results.rfe)
}

# Execute the function
rfe.results <- run.feature.selection(feature.vars = train.data[,-1], 
                                     class.var = train.data[,1])


# Visualize the results
rfe.results
varImp((rfe.results))
```


## Step 5 - Creating and Evaluating the First Version of the Model
 
 
```{r evaluating}
# Creating and Evaluating the Model
library(caret) 
library(ROCR) 

# Utility library for building graphics
source("plot_utils.R") 

## separate feature and class variables
test.feature.vars <- test.data[,-1]
test.class.var <- test.data[,1]

# Building a logistic regression model
formula.init <- "credit.rating ~ ."
formula.init <- as.formula(formula.init)
lr.model <- glm(formula = formula.init, data = train.data, family = "binomial")

# Visualize model
summary(lr.model)

# Testing the model using training data
lr.predictions <- predict(lr.model, test.data, type="response")
lr.predictions <- round(lr.predictions)

# Evaluating the model
confusionMatrix(table(data = lr.predictions, reference = test.class.var), positive = '1')
```


## Step 6 - Optimizing the model


```{r optimizing}
## Feature selection
formula <- "credit.rating ~ ."
formula <- as.formula(formula)
control <- trainControl(method = "repeatedcv", number = 10, repeats = 2)
model <- train(formula, data = train.data, method = "glm", trControl = control)
importance <- varImp(model, scale = FALSE)
plot(importance)


# Building the model with the select variables
formula.new <- "credit.rating ~ account.balance + credit.purpose + previous.credit.payment.status + savings + credit.duration.months"
formula.new <- as.formula(formula.new)
lr.model.new <- glm(formula = formula.new, data = train.data, family = "binomial")

# Visualizing the model
summary(lr.model.new)

# Testing the model with the test data
lr.predictions.new <- predict(lr.model.new, test.data, type="response") 
lr.predictions.new <- round(lr.predictions.new)

# Evaluating the model
confusionMatrix(table(data=lr.predictions.new, reference=test.class.var), positive='1')
```

## Step 7 - ROC Curve and Final Evaluating the model

```{r curve}
# Assessing model performance

# Creating ROC curves
lr.model.best <- lr.model
lr.prediction.values <- predict(lr.model.best, test.feature.vars, type = "response")
predictions <- prediction(lr.prediction.values, test.class.var)
par(mfrow = c(1,2))
plot.roc.curve(predictions, title.text = "ROC Curve")
plot.pr.curve(predictions, title.text = "Precision/Recall Curve")
```


